### 2023-08-18 Paper/paper.Rmd
Update paper with references and literature review

---
title: "The Gender Pay Gap in the General Social Survey"
author: "Alex Amaguaya, Lea Bergmann"
cols_authors: 3
subtitle: "Statistical Learning"
deadline: "tomorrow"
type: "Term Paper"
date: "today"
supervisors: "Prof. Dr. Thomas Deckers, Karolina Gliszczynska"
output:
  pdf_document:
    keep_tex: yes
    template: template.tex
    fig_caption: yes
    citation_package: biblatex
    number_sections: true
toc: true
lot: true
lof: true
graphics: true
biblio-title: References
fontsize: 11pt
geometry: lmargin=2.5cm,rmargin=2.5cm,tmargin=2.5cm,bmargin=2.5cm
biblio-files: references.bib
classoption: a4paper
---

<!-- % Template Version 1.1 -->

<!-- below function does some formatting for images; leave this untouched unless you know better :-) -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magick)
library(tidyverse)
library(stargazer)
crop <- function(im, left = 0, top = 0, right = 0, bottom = 0) {
  d <- dim(im[[1]]); w <- d[2]; h <- d[3]
  image_crop(im, glue::glue("{w-left-right}x{h-top-bottom}+{left}+{top}"))
}
```

# Introduction

In this paper we predict individuals' wages using data from the General Social Survey (GSS, 1974-2018). The variable of interest is `realrinc` which varies between 227\$ and 480,145\$\footnote{In constant 1986 USD.}. Hence, we can use regression techniques to predict individuals' wages. Additionally, we can consider whether `realrinc` varies by gender and if so, how much.

The data from the GSS allow us to consider various control variables such as gender, age, education, occupation, number of children and marital status.

The prediction of individuals' wages has been an issue in previous literature. As wages are relevant to companies they want to know how predict wages [@Chakraborti]. Some papers make use of classification techniques to predict wages [@Chakraborti]. As we have numeric values of real wages and not a categeorical variable we do not make use of classification techniques but prefer linear regression techniques such as OLS and Lasso regressions as well as random forests. We prefer these techniques as they handle categorical variables and outliers well, which are features of our data [@Cutler]. Wages were already predicted using random forests by [@Eichinger]. They find that usnig random forests results in better predictions than using linear models. Additionally, they show that gender barely effects the wage predictions.

Based on the previous literature, we use severall regression techniques to predict `realrinc`:

-   Linear Regression

-   Logistic Regression

-   Lasso, Ridge and Elastic Net Regression

-   Gradient Boosting

-   Random Forest

In order to evaluate which of the regression techniques predicts `realrinc` the best we split the data set into a train (70%) and a test (30%) dataset. We fit the aforementioned regression models on the train data and evaluate the root mean squared error (RMSE) using the validation data. We choose the RMSE as it is more sensitive to large errors [@Arour].

[Write something about cross validation?]

# Data Preparation

Before conducting the regression analysis we prepared the data. The provided dataset has 54,842 observations and for 38.55% of the observations the outcome variable `realrinc` is missing. As we cannot predict the wage for those observations we removed them. This left us with a dataset containing 33,702 observations. Additionally, we do not know the occupation for some observations (`occ10 = 9997` ) and removed those. This diminishes the dataset to 33,244 observations.

In this dataset still there were missing some values for some observations, e.g., number of children or occupation. To impute these values we used the `mice` function. We imputed age, occupational prestige and number of children with predictive mean matching. For the factor variables education and marital status we used polytomous logisitic regression. We receive six imputation datasets and use mean estimation for age, occupational prestige and number of children.

With this dataset we create dummy variables for the factor variables. Additionally, we create interaction variables.

[Something about ANOVA test for interaction variables].

[Should we consider outlier handling? cf. Eichinger 2022]

In a final step we calculate the correlation matrix using Spearman's rank correlation coefficient. We remove all variables that have a higher than 90% correlation with each other. Our final dataset then contains 33,244 observations of 361 variables.

[Something about how we define the x variables for the models]

# Model Training and Tuning

We used five different regression techniques to predict the individuals' wages using the final dataset. We fit every model on our training data using the `train` function from the `caret` package. To choose the best parameter combination we use cross-validation within all train functions.

## Linear and Logistic Regression

Our baseline linear model uses an OLS regression including all interaction variables. This results in an RMSE of 25,506.05.

``` r
formula <- as.formula(paste("realrinc ~", paste(x_cols_dummys, collapse = " + ")))
mod_full <- train(formula,
                  data = train, 
                  method = "lm",  
                  trControl = trainControl(method = "cv"))
```

Given some extreme outliers in the income distribution of our data (which is typical for wage data) we follow the economic literature [@Ermini] and fit a logistic model to our data, using the log of `realrinc` as our dependent variable. This results in a higher RMSE than the linear model (25,909.82).

``` r
formula2 <- as.formula(paste("log_realrinc ~", paste(x_cols_dummys, collapse = " + ")))
mod_loginc <- train(formula2,
                    data = train, 
                    method = "lm",  
                    trControl = trainControl(method = "cv"))
```

In both models the coefficient of female is statistically significant with p \< 0.001. It suggests that being female decreases the real income by approximately 10,192\$ in the linear model or by 40% in the logistic model.

## Subset Selection

As the dataset contains many variables (361) and even our selection of dummy variables for the right-hand side of the regression leaves us with 53 variables, we make use of regression techniques for subset selection. We use lasso, ridge and elastic net regressions. These methods allow us to fit a model that contains all variables. Then the model regularizes the coefficient estimates and shrinks them towards zero. In effect, we reduce the estimates' variance which improves the model fit.

We start with a ridge regression.

$$
RSS + \lambda \sum^p _{j=1} \beta^2_j 
$$

Hence, the coefficient estimates in a ridge regression minimize the sum of the residual sum of squares (RSS) and the sum of squared coefficients multiplied with a tuning parameter $\lambda$. We estimate $\lambda$ using cross-validation.

``` r
ridge <- train(formula, data = train,
  method = "glmnet", trControl = trainControl(method = "cv"),
  tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) 
```

Perform a ridge regression to predict real incomes yields an RMSE of 25,524.85.

In contrast to other subset selection methods ridge regression does not exclude any variables from the regression model. Hence, the model is still fitted on 53 covariates. To overcome this issue we also use a lasso regression.

$$
RSS + \lambda \sum^p_{j=1} | \beta_j |
$$

In contrast to the ridge regression the lasso regression forces some of the coefficient estimates to be equal to zero if $\lambda$ is sufficiently large. Hence, the regression is not performed on all covariates.

``` r
lasso <- train(formula, data = train, 
       method = "glmnet", trControl = trainControl(method = "cv"), 
       tuneGrid = expand.grid(alpha = 1, lambda = seq(0,500,1)))
```

To perform the lasso regression in `R` we set $\alpha$ equal to one. The lasso regression yields a lower RMSE than the ridge regression (25,504.58 vs. 25,524.85) and even a lower RMSE than the linear regression.

In a final step we use elastic net regularization which is a linear combination of ridge and lasso regression. The elastic net regularization allows $\alpha$ to vary between zero and one. This results in an RMSE of 25,241.7.

``` r
elasticnet <- train(formula, data = train,
       method = "glmnet", trControl = trainControl(method = "cv"),
       tuneGrid = expand.grid(alpha = seq(from=0, to=1, by = 0.1),
       lambda = seq(from=0, to=0.15, by = 0.001)))
```